---
title: "HW05 Part 1"
author: "Olivia Pura"
date: "8/3/2020"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW05 Part 1

In Part 1 of HW05, we're tidying up the dad_mom.txt dataset. To do this, we have to load the tidyverse package. I'm also loading in the knitr package to utilize the kable() function.

```{r packages, message=FALSE}
library(tidyverse)
library(knitr)
```

Now, we're reading in the dad_mom.txt dataset and printing it using kable().

```{r file, echo=FALSE}
dad_mom <- read.table("dad_mom.txt", header = TRUE)

kable(dad_mom, format = "markdown", align = "l")
```

In order for this data to be tidy, our columns must represent the four variables - family, parent, name, and income. 

```{r tidying, echo=FALSE}
tidy_dad_mom <- dad_mom %>%
#uniting data for the dads and moms separately
  unite(Dad, name_dad, income_dad) %>%
  unite(Mom, name_mom, income_mom) %>%
  
#gathering the data 
  gather(key = "Parent", value = "name_income", -fam_id) %>%
  
#separating the name_income column to create the final two columns
  separate(name_income, into = c("Name", "Income")) %>%
  
#renaming the fam_id column for cleanliness
  rename(Family = fam_id)
```
 
Now that the data is tidy, we can view it as a table using kable()

```{r final, echo=FALSE}
kable(tidy_dad_mom, format = "markdown", align = "l")
```