---
title: "HW05"
author: "Xing Yanfeng"
date: "2020/8/2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 1

### Load libraries and read in data.

```{r libraries data}
# load libraries
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
# read in data
dad_mom <- read_delim("dad_mom.txt", delim = "\t")
```

### Tidying data

```{r tidying}
tidy_dad_mom <- dad_mom %>% 
  gather(key = "key", value = "value", 2:5) %>% 
  separate(col = key, into = c("key", "relationship")) %>% 
  spread(key = key, value = value, convert = T) %>% 
# rearrange columns to make it more comfortable
  select(fam_id, relationship, name, income)
print(tidy_dad_mom)
```

## Assignment 2

### Read in data

```{r read in data}
AUC_1 <- read_csv("CTRP_files/AUC_1.csv")
AUC_2 <- read_csv("CTRP_files/AUC_2.csv")
cell_line <- read_csv("CTRP_files/cancer_cell_line_info.csv")
compound <- read_csv("CTRP_files/compound_info.csv")
experiment <-  read_csv("CTRP_files/Experiment_info.csv")
```

### Join 5 csv files

```{r join 5 csv files}
cancer_auc <- AUC_1 %>%
# bind AUC_1 and AUC_2
  bind_rows(AUC_2)  %>%
# add experiment information
  inner_join(unique(experiment), by = c("experiment_id" = "expt_id"))  %>%
# add compound information
  left_join(compound, by = "master_cpd_id") %>% 
# add cell line information
  left_join(cell_line, by = "master_ccl_id")
```

### Question 1

Which cancer type has the lowest AUC values to the compound "vorinostat"?

```{r question 1 data}
vorinostat <- cancer_auc %>% 
# filter AUC values and cancer type of vorinostat treated experiments
  select(area_under_curve, cpd_name, cancer_type) %>% 
  filter(cpd_name == "vorinostat") %>% 
  drop_na(cancer_type) %>% 
# calculate average AUC values
  group_by(cancer_type) %>% 
  summarize(AUC = round(mean(area_under_curve), digits = 2)) %>% 
  arrange(AUC) %>% 
  mutate(cancer_type = factor(cancer_type, levels = cancer_type))
```

```{r question 1 graph}
# plot
ggplot(vorinostat, aes(x = cancer_type, y = AUC)) +
  geom_col(fill = "#00CDFF") +
# add labels
  labs(x = "Cancer Type", 
       y = "Area Under Curve", 
       title = "AUC of Different Cancer Types Treated by Vorinostat") +
  geom_text(aes(label = AUC), vjust = -0.5, size = 2.5) +
# adjust theme
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = -30, hjust = 0)) 
```

```{r}
print(paste("The cancer type having the lowest AUC values to the compound vorinostat is ", gsub(as.character(vorinostat$cancer_type)[1], pattern = "_", replacement = " "), ".", sep = ""))
```

### Question 2

Which compound is the prostate cancer cell line 22RV1 most sensitive to?

```{r question 2 data}
prostate_cancer <- cancer_auc %>% 
  select(area_under_curve, cpd_name, ccl_name) %>% 
  filter(ccl_name == "22RV1") 
prostate_cancer
```

There are no observations with `ccl_name == "22RV1"` in `cancer_auc`, and there might be some problems with the dataset. So I answered the question "Which compound is the breast cancer cell line KPL1 most sensitive to?" instead.

```{r question 2 data2}
KPL1 <- cancer_auc %>% 
# filter AUC values of KPL1 treated by different compounds
  select(area_under_curve, cpd_name, ccl_name) %>% 
  filter(ccl_name == "KPL1") %>% 
# calculate average AUC values of each compound
  group_by(cpd_name) %>% 
  summarize(AUC = round(mean(area_under_curve), digits = 2)) %>% 
  arrange(AUC) 
```

```{r question 2 answer}
print(paste("The breast cancer cell line KPL1 is most sensitive to ", KPL1[1, 1], ".", sep = ""))
```


### Question 3

For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

```{r question 3 data}
EGFR <- cancer_auc %>% 
# filter AUC values of breast cancer cell lines treated by EGFR targeting compounds
  select(area_under_curve, cpd_name, gene_symbol_of_protein_target, ccl_name, cancer_type) %>% 
  filter(cancer_type == "breast", grepl("EGFR", gene_symbol_of_protein_target)) %>% 
# calculate average AUC values of each cell line
  group_by(cpd_name, ccl_name) %>% 
  summarize(AUC_ccl = mean(area_under_curve)) %>% 
# calculate average AUC values of each compound
  summarize(AUC = round(mean(AUC_ccl), digits = 2)) %>% 
  arrange(AUC) %>% 
  mutate(cpd_name = factor(cpd_name, levels = cpd_name))
```

```{r question 3 graph}
# plot
ggplot(EGFR, aes(x = cpd_name, y = AUC)) +
  geom_col(fill = "#00CDFF") +
# add labels
  labs(x = "Compound", 
       y = "Area Under Curve", 
       title = "AUC of Breast Cancer Cell Lines Treated by EGFR Targeting Compounds") +
  geom_text(aes(label = AUC), vjust = -0.5, size = 2.5) +
# adjust theme
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = -30, hjust = 0))
```

```{r question 3 answer}
print(paste("For the 10 compounds that target EGFR, ", as.character(EGFR$cpd_name)[1], " has the lowest AUC values in the breast cancer cell lines.", sep = ""))
```

```{r session info}
devtools::session_info()
```

