---
title: "HW05_Julia"
author: "Julia Shangguan"
date: "8/1/2020"
output: github_document 
---

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

Loading in necessary packages:
```{r load libraries}
#install.packages('tidyverse')
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Part 1: Tidying the dad_mom file

```{r tidy dad_mom data}
parent_data <- read.delim("dad_mom.txt") #read dataframe

parent_data_tidy <- parent_data %>%
  gather(key = "parent", value= "name", name_dad, name_mom) %>% #create name column
  gather(key= "key1", value = "income", income_dad, income_mom) %>% #create income column
  select(-key1)

parent_data_tidy$parent <- recode(parent_data_tidy$parent, "name_dad" = "dad", "name_mom" = "mom") #create parent's gender column

parent_data_tidy
```

## Part 2: Joining together CTRP data

```{r CTRP data}
#read all dataframes
cancer_cell_line <- read_csv("CTRP_files/cancer_cell_line_info.csv")
AUC_1 <- read_csv("CTRP_files/AUC_1.csv")
AUC_2 <- read_csv("CTRP_files/AUC_2.csv")
compound_info <- read_csv("CTRP_files/compound_info.csv")
experiment_info <- read_csv("CTRP_files/Experiment_info.csv")
```

```{r}
#looking at each dataframe
glimpse(cancer_cell_line)
glimpse(AUC_1)
glimpse(AUC_2)
glimpse(compound_info)
glimpse(experiment_info)
```
```{r joining the tables}
#is it wrong to use full_join to merge all the tables? or should I be using inner_join?

AUC_all_compound<- AUC_1 %>%
  full_join(AUC_2, by = c("experiment_id", "area_under_curve", "master_cpd_id")) %>%
  full_join(compound_info, by = "master_cpd_id")

ccl_experiment_AUC_compound <- cancer_cell_line %>%
  full_join(experiment_info, by = "master_ccl_id") %>%
  full_join(AUC_all_compound, by=c("expt_id"="experiment_id"))

glimpse(ccl_experiment_AUC_compound)
```

Which cancer type has the lowest AUC values to the compound "vorinostat"? 

* answer = haematopoietic and lymphoid tissue
```{r}
vorinostat_auc <- ccl_experiment_AUC_compound %>%
  filter(cpd_name == "vorinostat") %>%
  arrange(area_under_curve)

ggplot(vorinostat_auc, aes(x= reorder(cancer_type, area_under_curve), y = area_under_curve)) + 
  geom_boxplot() +
 theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  scale_x_discrete(name = "cancer type")

```

Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)

* answer = ? For some reason, I have NA in the compound column

```{r}
ccl_experiment_AUC_compound %>%
  filter(ccl_name == "22RV1") %>%
  arrange(desc(area_under_curve))
```

For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

* answer = afatinib


```{r}

mean_auc <- ccl_experiment_AUC_compound %>%
  group_by(cpd_name, gene_symbol_of_protein_target) %>%
  filter(cancer_type == "breast") %>%
  filter(str_detect(gene_symbol_of_protein_target, "EGFR")) %>%
  summarize(mean_AUC = mean(area_under_curve)) %>%
  arrange(mean_AUC)

ggplot(mean_auc, aes(x= reorder(cpd_name, mean_AUC), y =mean_AUC)) + 
  geom_col() +
  scale_x_discrete(name = "compound name")

```
