---
title: "HW06"
author: "Brock"
date: "8/3/2020"
output: html_document
---
#Goal: Tidy and join data :)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse") #Tidyverse functionalities
library("lubridate") #Better data/time parsing
library("knitr") #Nice outputs for tibbles/ dataframes
library("here") #Pathing for importing data
```

Data importing is made easier with the here() function from the "here" package.

```{r Data import, message = FALSE}
dad_mom <- read_delim("dad_mom.txt", delim = "\t")

auc_1 <- read_csv(here("./CTRP_files/AUC_1.csv")) #here() outputs path of current project directory; improved readability vs "C:/BLAH/BLAH/BLAH/BLAH/etc"

auc_2 <- read_csv(here("./CTRP_files/AUC_2.csv"))

cancer_cell_line_info <- read_csv(here("./CTRP_files/cancer_cell_line_info.csv"))

compound_info <- read_csv(here("./CTRP_files/compound_info.csv"))

experiment_info <- read_csv(here("./CTRP_files/Experiment_info.csv")) #here() is MVP
```

I was stuck on this for quite awhile. I wanted to find a way to combine the name_ income_ variables into a single column, but I couldn't think of a great way to do that. I look forward to reviewing the code for this assignment to see what people came up with!

```{r tidy family data}
tidy_family <- dad_mom %>%
  gather(key = "key", value = "value", -fam_id, -name_dad, -name_mom) %>% #Gathers income and names of parents, keeps fam_id separate
  mutate("parent_name" = c(name_dad[1:3], name_mom[1:3])) %>% #I couldn't think of the better way to do this
  transmute(fam_id, parent_name, "parent_type" = str_replace(key, ".+_", ""), "income" = value)
```

The following are variables of each dataframe:
auc_1: experiment_id, area_under_curve, master_cpd_id and 
auc_2 as auc_1
cancer_cell_line_info: master_ccl_id, ccl_name, cancer_type
compound_info: master_cpd_id, cpd_name, gene_symbol_of_protein_target
experiment_info: expt_id, master_ccl_id, experiment_date, cells_per_well

Variable Links:
compound_info <=> auc_1 = auc_2 <=> experiment_info <=> cancer_cell_line_info

```{r joining CTRP_files}
ctrp <- auc_1 %>%
  full_join(auc_2, by = c("experiment_id", "area_under_curve", "master_cpd_id")) %>%
  full_join(compound_info, by = "master_cpd_id") %>%
  full_join(experiment_info, by = c("experiment_id" = "expt_id")) %>%
  full_join(cancer_cell_line_info, by = "master_ccl_id") #Decided to full_join at each step for maximum amount of data

```

```{r session info}
devtools::session_info()
```