---
title: "HW05"
output: github_document
author: Bei Wang
---

```{r load pkgs, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("magrittr")
library("knitr")
library("readr")
```

## 1. Tidying the dad_mom file

    Each variable must have its own column.
    Each observation must have its own row.
    Each value must have its own cell.

```{r dad_mom}
#read dataset
dad_mom<- read.table("dad_mom.txt", header=TRUE) 

#reshape
tidy_dad_mom<- dad_mom %>%
  unite("dad", name_dad, income_dad) %>%
  unite("mom", name_mom, income_mom) %>%
  pivot_longer(c(dad, mom), names_to="parent", values_to="income") %>%
  separate(income, c("name", "income"))

#print
tidy_dad_mom
```

## 2. Joining together CTRP data

```{r totalauc}
#read files
AUC_1 <- read_csv("CTRP_files/AUC_1.csv")
AUC_2 <- read_csv("CTRP_files/AUC_2.csv")
cancer_cell_line_info <- read_csv("CTRP_files/cancer_cell_line_info.csv")
compound_info <- read_csv("CTRP_files/compound_info.csv")
Experiment_info <- read_csv("CTRP_files/Experiment_info.csv")

#glimpse
glimpse(AUC_1)
glimpse(AUC_2)
glimpse(cancer_cell_line_info)
glimpse(compound_info)
glimpse(Experiment_info)

#Since AUC_1 and AUC_2 look like two parrelle data frame with links to the other files, so i join AUC_1 and _2 first, and then combin them with the other files

#join AUC_1 with _2
AUC<- AUC_1 %>%
  bind_rows(AUC_2)

#join auc with other files
total_AUC<- AUC %>%
  inner_join(Experiment_info, by=c("experiment_id"="expt_id")) %>%
  inner_join(compound_info, by="master_cpd_id") %>%
  inner_join(cancer_cell_line_info, by="master_ccl_id") 
```

* Which cancer type has the lowest AUC values to the compound "vorinostat"?

  * answer: haematopoietic_and_lymphoid_tissue
  
```{r q1}
#average AUC of different cancer types
q1<- total_AUC %>%
  select(area_under_curve, cancer_type, cpd_name) %>%
  filter(cpd_name== "vorinostat") %>%
  group_by(cancer_type) %>%
  drop_na(cancer_type) %>%
  summarize(AUC= mean(area_under_curve))
  
#plot
q1 %>%
  ggplot(aes(x=cancer_type, y=AUC, fill=cancer_type))+
  geom_col()+
  labs(title="Which cancer type has the lowest AUC values to the compound vorinostat", x=element_blank(), y="Average Area Under the Curve")+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  theme(legend.position="none")
```

* Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)

  * answer: LBH-589
  
```{r q2}
#it returns 0 row when using 22RV1 for ccl_name, unsure why.
total_AUC %>%
  select(ccl_name, cpd_name, area_under_curve) %>%
  filter(ccl_name=="22RV1") %>%
  arrange(area_under_curve)

#instead use prostate as in cancer_type 
q2<- total_AUC %>%
  select(cancer_type, cpd_name, area_under_curve) %>%
  filter(cancer_type=="prostate") %>%
  arrange(area_under_curve)

#select the lowest 5 in terms of auc for plotting
low5<- q2[1:5,] 

low5 %>%
  ggplot(aes(x=cpd_name, y=area_under_curve, fill= cpd_name)) +
  geom_col()+
  labs(title="For prostate cancer cell line, the five componds have the lowest AUC value", x=element_blank(), y="Area Under the Curve")+
  theme(legend.position="none")
```

* For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer

  * "WZ8040" has the lowest average value
  * There are only 3 compounds I filtered based on the question, instead of 10, I'm not sure if the way I did it is incorrect.  

```{r q3}
#filter the AUC according to the question
q3<- total_AUC %>%
  select(gene_symbol_of_protein_target, area_under_curve, cpd_name, cancer_type) %>%
  filter(gene_symbol_of_protein_target=="EGFR", cancer_type=="breast") %>%
  group_by(cpd_name) %>%
  summarize(AUC=mean(area_under_curve)) #my filtered cpd_name has only 3 different values instead of 10?

#plot
q3 %>% ggplot(aes(x=cpd_name, y=AUC, fill= cpd_name)) +
  geom_col() +
  labs(title="AUC values of compounds that target EGFR in breast cancer", x=element_blank(), y="Average Area Under the Curve")+
  theme(legend.position="none")
```