---
title: "HW05"
output: github_document
author: "Daniel Krochmal"
date: "8/4/2020"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r load libraries, echo = FALSE, message = FALSE}
library(readr)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
```


## Part 1


``` {r import data, echo = FALSE, message = FALSE}
Dad_mom <- read_tsv("dad_mom.txt") #Importing data using readr
```

``` {r tidy data}

#I will use a sequence of gather(), separate() and spread() verbs to clean-up the datafame

tidyDad_mom <- Dad_mom %>%
  
  gather(key = "var", value = "value",
         convert = FALSE, -fam_id) %>%
  
  #Since the "var" column contains two variables, I'm separating them into "var" (name,income) and "relation" (mom,dad)
  separate(col = var, into = c("var", "relation")) %>%
  
  #Since name and income should have their own column, I'm using spread() to accomplish that 
  spread(key = var, value = value, convert = TRUE) %>%
  
  #Reordering column using select. I wanted the "name" to be a second column, so for the rest I used everything()
  select(fam_id, name, everything()) %>%
  rename(family = fam_id)

kable(
  tidyDad_mom,
  format = "markdown",
  align = "c"
)
```
## Part 2

``` {r import files, echo = FALSE, message = FALSE}

#importing csv files using readr
auc1 <- read_csv("CTRP_files/AUC_1.csv")
auc2 <- read_csv("CTRP_files/AUC_2.csv")
ccl <- read_csv("CTRP_files/cancer_cell_line_info.csv")
cpd <- read_csv("CTRP_files/compound_info.csv")
exp <- read_csv("CTRP_files/Experiment_info.csv")

```
``` {r join data, echo = FALSE}
ctrp_data <- auc1 %>%
  
  #Binding second part of the experimental data (auc2) to auc1
  bind_rows(auc2) %>%
  
  #Adding experiment info to the experimental data
  left_join(exp, by = c("experiment_id" = "expt_id")) %>%
  
  #Adding compound info
  left_join(cpd, by = "master_cpd_id" ) %>%
  
  #Adding cell line info
  left_join(ccl, by = "master_ccl_id")

```

### Question 1
Which cancer type has the lowest AUC values to the compound "vorinostat"?

``` {r Q1, echo = FALSE}

#Selecting for the variables relevant to the question, filtering for the compound of interest and grouping by cancer type
vorinostat <- ctrp_data %>%
  select(area_under_curve, cpd_name, cancer_type) %>%
  filter(cpd_name == "vorinostat") %>%
  drop_na(cancer_type) %>%
  group_by(cancer_type)

#To get the answer in a simple way I'm using a boxplot with x axis in ascending AUC order
ggplot(vorinostat, mapping = aes(x = reorder(cancer_type, area_under_curve), y = area_under_curve)) +
  geom_boxplot() +
  labs(
    x = "Cancer type",
    y = "AUC"
  ) +
  
  theme_classic() +
  
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 8))

```

**Autonomic ganglia displayed the lowest AUV values to vorinostat**

### Question 2
Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)


``` {r Q2, echo = FALSE, message = FALSE}

data_22rv1 <- ctrp_data %>%
  select(area_under_curve, cpd_name, ccl_name) %>%
  filter(ccl_name == "22RV1") %>%
  group_by(cpd_name) %>%
  summarize(AUC_22rv1 = mean(area_under_curve)) %>%
  arrange(AUC_22rv1)

#Since the resulting dataframe is too large to plot it reasonably, I'm printing the answer using kable().
kable(
  head(data_22rv1, n = 1),
  format = "markdown",
  col.names = c("Compound name", "AUC")
)

```

**Prostate cancer cell line 22RV1 was the most sensitive to leptomycin B**

### Question 3
For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

``` {r Q3, echo = FALSE}

egfr_data <- ctrp_data %>%
  select(area_under_curve, cancer_type, cpd_name, gene_symbol_of_protein_target) %>%
  
  #Using grepl() to filter for the values in gene_symbol_of_protein_target that contain specified string ("EGFR")
  filter(grepl("EGFR", gene_symbol_of_protein_target), cancer_type == "breast") %>%
  group_by(cpd_name)

ggplot(egfr_data, aes(x = reorder(cpd_name, area_under_curve), y = area_under_curve)) +
  geom_boxplot() +
  labs(
    x = "Compound name",
    y = "AUC"
  ) +
  
  theme_classic() +
  
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 8))

```

**Out of EGFR-targeting compounds, Afatinib had the lowest AUC values in breast cancer cell lines.**



```{r session info, echo = FALSE}
devtools::session_info()
```