---
title: "HW05: mom_dad and CTRP data"
author: "Laura N. Cespedes Arias"
date: "31/7/2020"
output: github_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and importing data
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)    
library(knitr)
```

# 1) The mom_dad data 

## Importing the data
```{r include=TRUE}
# read in the data
dad_mom <- read.table("dad_mom.txt",h=T)

#These is how the table looked originally:
knitr::kable(dad_mom)
```

## Tidying the dad_mom file

Pretty sure there is a shorter and cleaner way to do this, but this is was the best solution I could think of:

First, I used pivot_longer to reduce the number of columns (now each mom and dad per family have their own row). I created a column named "parent" where I specified if the person was a mom or a dad. Then I replaced the values of these variable (using mutate & replace). After that, I created a new column called "income" which has the value for the income of the corresponding parent (extracting values from income_dad and income_mom depending of the "parent" column) For this I used mutate & if_else. Lastly, I eliminated the income_dad and income_mom columns.

```{r include=TRUE}
 dad_mom_v2 <-
  dad_mom %>% pivot_longer(c(name_dad, name_mom),
                           names_to = "parent",
                           values_to = "name") %>% mutate(parent = replace(parent, parent == "name_mom", "mom")) %>% mutate(parent = replace(parent, parent == "name_dad", "dad")) %>% mutate(income = if_else(parent == "mom", income_mom, income_dad)) %>% select(-c(income_dad, income_mom))

#This is how the table looks after tidying!
knitr::kable(dad_mom_v2)
```

# 1) The CTRP data

## Importing the data

```{r include=TRUE}
# I tried to get all files automatically using apply but I failed to change the working directory within the project :/. So I ended up doing it manually:
AUC_1 <- read.csv("./CTRP_files/AUC_1.csv",h=T)
AUC_2 <- read.csv("./CTRP_files/AUC_2.csv",h=T)
Experiment_info <- read.csv("./CTRP_files/Experiment_info.csv",h=T)
cancer_cell_line_info <- read.csv("./CTRP_files/cancer_cell_line_info.csv",h=T)
compound_info <- read.csv("./CTRP_files/compound_info.csv",h=T)
```

## Joining together CTRP data

First, I combined the AUC tables (which have the same columns, so I used bind_rows).

```{r include=TRUE}
AUC <- bind_rows(AUC_1,AUC_2)
knitr::kable(head(AUC))
```

Then, I merged the AUC data with the other data frames.

```{r include=TRUE}

#First I changed the column name of the experiment ID in the "AUC" table so it matched the one from the "Experiment_info" table.
AUC_2 <- AUC %>% rename(expt_id=experiment_id) 

#Then I merged the data tables based on the common column (I did this subsequently starting from AUC + Experimenta_info, which had in common the column (i.e. variable) "expt_id")
AUC_experiment <- full_join(AUC_2,Experiment_info,by = "expt_id")
AUC_experiment_cancer <- full_join(AUC_experiment,cancer_cell_line_info,by = "master_ccl_id")
AUC_experiment_cancer_compound <- full_join(AUC_experiment_cancer,compound_info,by = "master_cpd_id")

#Just to check the final table column names
colnames(AUC_experiment_cancer_compound)

#Exported the new data frame, just in case.
write.csv(AUC_experiment_cancer_compound,"CTRP_data_all.csv")
```

## Answers to questions

### Which cancer type has the lowest AUC values to the compound "vorinostat"?

Based on the figure **hematopoietic and lymphoid tissue cancer** has the lowest AUC values on average. However, the lowest AUC value correspond to **upper aerodigestive tract cancer**.

```{r include=TRUE}
#First I filtered the complete date set to include only the observations involving the compound "vorinostat"
vorinostat_data <- AUC_experiment_cancer_compound %>% filter(cpd_name=="vorinostat")

#Then, I built a boxplot to visualize the difference in AUC among cancer types:
ggplot(vorinostat_data,mapping = aes(x = cancer_type, y= area_under_curve)) +
  geom_jitter(col="steelblue4",size=0.8)+
  geom_boxplot(fill="steelblue4",alpha=0.6,outlier.colour = NA)+
  theme_bw()+
  theme(axis.title = element_text(color="black", size=12))+
  xlab("Cancer type") +
  ylab("Area under curve (AUC)")+
  theme(axis.text.x = element_text(color="black", size=8, angle=45,hjust=1))
```

### Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)

**leptomycin B** is the compound to which the prostate cancer cell line 22RV1 is most sensitive to.
```{r include=TRUE}
#First I filtered the complete date set to include only the observations involving prostate cancer cell line "22RV1"
line22RV1_data <- AUC_experiment_cancer_compound %>% filter(ccl_name=="22RV1") %>% filter(cancer_type=="prostate")

#top_n works to select only the top values for a variable in a data frame (and - is used to select the minimum value instead)
(top_n(line22RV1_data,-1,area_under_curve))$cpd_name

```

### For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

The coumpound with lowest AUC values in average is **afatinib**. However, I identified 11 compounds that target EGFR, not 10.

```{r include=TRUE}
#First I filtered the complete date set to have one including only compounds that target EGFR 
#Because there can be multiple values in the "gene_symbol_of_protein_target" (separated by ;), I used "grep" instead of "=="
#I also filtered to only include breast cancer data.
EGFR_data <- AUC_experiment_cancer_compound %>% filter(grepl("EGFR",gene_symbol_of_protein_target)) %>%filter(cancer_type=="breast")

#Then, I built a boxplot to visualize the difference in AUC among compounds:
ggplot(EGFR_data,mapping = aes(x = cpd_name, y= area_under_curve)) +
  geom_jitter(col="seagreen4",size=2)+
  geom_boxplot(fill="seagreen4",alpha=0.6,outlier.colour = NA)+
  theme_bw()+
  theme(axis.title = element_text(color="black", size=12))+
  xlab("Compound name") +
  ylab("Area under curve (AUC)")+
  theme(axis.text.x = element_text(color="black", size=8, angle=45,hjust=1))
```

## Session info
```{r}
# always good to have this for reproducibility purposes
devtools::session_info()
```

