---
title: "HW05"
author: "Ricardo Muñiz Trejo"
date: "2/8/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
```

## Tidying the dad_mom file

Untidied version of the dad_mom data frame:


```{r}
dad_mom_messy <- read_tsv("dad_mom.txt")

kable(dad_mom_messy, format = "markdown")
```

Tidied version of the dad_mom data frame:

```{r}

#Ok, this one can be confusing. The objective is to have 4 variables: fam_id, name, parent, income

dad_mom_tidy <- dad_mom_messy %>%
  #First we need to create the parent variable which considers if its the mom or the dad
  separate(col = name_dad, into = c("name_dad", "parent_dad")) %>%
  separate(col = name_mom, into = c("name_mom", "parent_mom")) %>%
  replace_na(list(parent_dad = "dad",
                  parent_mom = "mom")) %>%
  
  #For the moment, I joined the name and parent variables into one
  unite(name_dad, parent_dad, col = "name_dad", sep = "_") %>%
  unite(name_mom, parent_mom, col = "name_mom", sep = "_") %>%
  
  #Now we need to spread the combined name variables to gain columns with the parents names
  spread(key = name_dad, value = income_dad) %>%
  spread(key = name_mom, value = income_mom) %>%
  
  #We gather everything, eliminating NAs, to get a single column
  gather(key = "name", value = "income", -fam_id, na.rm = TRUE) %>%
  
  #Finally, we separate name and parent variables to finish
  separate(col = name, into = c("name", "parent")) %>%
  arrange(fam_id)
  
kable(dad_mom_tidy, format = "markdown")

```

## Joining together CTRP data

#### Reading and importing files:

```{r}

#Let's make data frames for every file in CTPR_files

AUC_1 <- read_csv("CTRP_files/AUC_1.csv")
AUC_2 <- read_csv("CTRP_files/AUC_2.csv")
cancer_cell_line_info <- read_csv("CTRP_files/cancer_cell_line_info.csv")
compound_info <- read_csv("CTRP_files/compound_info.csv")
experiment_info <- read_csv("CTRP_files/experiment_info.csv")


```

#### Joining process:

```{r}

#Bind the AUC files to get a master file
AUC_total <- AUC_1 %>%
  bind_rows(AUC_2)

glimpse(AUC_total)

CTRP_df <- AUC_total %>%
  inner_join(experiment_info, by = c("experiment_id" = "master_ccl_id")) %>%
  inner_join(cancer_cell_line_info, by = c("experiment_id" = "master_ccl_id")) %>%
  inner_join(compound_info, by = "master_cpd_id") %>%
  drop_na(gene_symbol_of_protein_target) %>%
  separate_rows(gene_symbol_of_protein_target, sep = ";")


glimpse(CTRP_df)

```

#### Answering the questions:


Once those files are all together, you should have no problem answering the following questions (with graphs):

* Which cancer type has the lowest AUC values to the compound "vorinostat"?

```{r}

theme_set(theme_bw())

CTRP_df_q1 <- CTRP_df %>%
  select(cpd_name, cancer_type, area_under_curve) %>%
  filter(cpd_name == "vorinostat") %>%
  group_by(cancer_type) %>%
  summarize(AUC = mean(area_under_curve)) %>%
  arrange(AUC)

CTRP_df_q1

ggplot(CTRP_df_q1, aes(cancer_type, AUC)) +
  geom_col() +
  coord_flip()
  
```



* Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)

```{r}

CTRP_df_q2 <- CTRP_df %>%
  select(cancer_type, ccl_name, cpd_name, area_under_curve) %>%
  filter(cancer_type == "prostate", ccl_name == "22RV1") %>%
  group_by(cpd_name) %>%
  summarize(AUC = mean(area_under_curve)) %>%
  arrange(AUC)


CTRP_df_q2

```



* For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?


```{r}

CTRP_df_q3 <- CTRP_df %>%
  filter(gene_symbol_of_protein_target == "EGFR", cancer_type == "breast") %>%
  group_by(cpd_name) %>%
  summarize(AUC = mean(area_under_curve)) %>%
  arrange(AUC)

CTRP_df_q3

```



## Session info

```{r}
devtools::session_info()
```

