---
title: "HW05 Parts 1 and 2"
author: "Luke Dyer"
date: "8/3/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1
This cleans up the dad_mom.txt file and makes it into a manageable table.

```{r dad_mom}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)

dad_mom_data <- read.delim("dad_mom.txt")

tidy_dad_mom <- dad_mom_data%>%
  unite(Dad, name_dad, income_dad)%>%
  unite(Mom, name_mom, income_mom)%>%
  gather(key="Parent", value="name_income", -fam_id)%>%
  separate(name_income, into=c("Name", "Income"))%>%
  arrange(fam_id)%>%
  rename(Family_Number = fam_id)

kable(tidy_dad_mom)

```

##Part 2, Combining Data Sources

```{r cancer_data}

AUC_1 <- read.csv("CTRP_files/AUC_1.csv")
AUC_2 <- read.csv("CTRP_files/AUC_2.csv")
cancer_cell_line_invo <- read.csv("CTRP_files/cancer_cell_line_info.csv")
compound_info <- read.csv("CTRP_files/compound_info.csv")
experiment_info <- read.csv("CTRP_files/experiment_info.csv")

cancer_data <- AUC_1%>%
  full_join(AUC_2, by=c("experiment_id", "area_under_curve", "master_cpd_id"))%>%
  full_join(compound_info, by="master_cpd_id")%>%
  full_join(experiment_info, by=c("experiment_id"="expt_id"))%>%
  full_join(cancer_cell_line_invo, by="master_ccl_id")

```

##Part 2, Question 1
The below graph summarizes the Effect of Vorinostat on different cancer types. You can see haematopoietic and lymphoid tissue has the lowest average AUC, while the lowest extreme comes from tissue in the upper aerodigestive tract.

```{r question1}
low_AUC <- cancer_data%>%
  select(experiment_id, cpd_name, cancer_type, area_under_curve)%>%
  filter(cpd_name=="vorinostat")%>%
  arrange(area_under_curve)%>%
  group_by(cancer_type)

ggplot(low_AUC, aes(x=reorder(cancer_type, area_under_curve,), y=area_under_curve))+
geom_boxplot()+
  theme(axis.text.x=element_text(angle=65, hjust=0.9))+
  labs(x="Cancer", y="Distribution", title="Vorinostat Effect on Cancer Types")

```

##Part 2, Question 2
In this case, only one row of data was required to answer the question. Therefore, a simple kable table was used to show that the cancer cell line 22RV1 was most sensitve to compound leptomycin B.

```{r question2}
prostate_sens <- cancer_data%>%
  select(experiment_id, area_under_curve, cpd_name, ccl_name, cancer_type)%>%
  filter(ccl_name == "22RV1")%>%
  arrange(area_under_curve)%>%
  top_n(n=-1, wt = area_under_curve)%>%
  rename(Experiment=experiment_id, AUC=area_under_curve, Compound=cpd_name, Cell_Line=ccl_name, Cancer_Type=cancer_type)

kable(prostate_sens)
```

##Part 2, Question 3
Based on the graph, the compound WZ8040 had the lowest AUC values in breast cancer cell lines.

```{r question3}
EGFR_AUC <- cancer_data%>%
  select(experiment_id, gene_symbol_of_protein_target, cpd_name, area_under_curve, cancer_type)%>%
  filter(gene_symbol_of_protein_target=="EGFR", cancer_type=="breast")%>%
  group_by(cpd_name)%>%
  summarize(average_AUC=mean(area_under_curve))

ggplot(EGFR_AUC, aes(x=cpd_name, y=average_AUC))+
  geom_col()+
  labs(x="Compound Name", y="Average AUC", title="Sensitivity of Breast Cancer Line EGFR to Compounds")


```

