---
title: "Joining together CTRP data"
output: github_document
---

## 1. Load packages
```{r load packages}

library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(knitr)
library(lubridate)

```

## 2. Load data
```{r load data}

setwd("~/Desktop/GitHub/Homework_5/HW05/CTRP_files")

AUC_1 <- fread("AUC_1.csv")
AUC_2 <- fread("AUC_2.csv")
experiment_info <- fread("Experiment_info.csv")
cancer_cell_line <- fread("cancer_cell_line_info.csv")
compound <- fread("compound_info.csv")
  
```

## 3. Tidy the data frames

### 3a. AUC_1
```{r AUC_1}

# Rename the columns in AUC_1 file

AUC_1_rename <- AUC_1 %>%
    rename("Exp._ID" = experiment_id, "AUC" = area_under_curve, "Compound_ID" = master_cpd_id) %>%
    select("Exp._ID", "Compound_ID", "AUC")

table1 <- head(AUC_1_rename)
kable(table1, align = "ccc")

```


### 3b. AUC_2
```{r AUC_2}

# Rename the columns in AUC_2 file

AUC_2_rename <- AUC_2 %>%
    rename("Exp._ID" = experiment_id, "AUC" = area_under_curve, "Compound_ID" = master_cpd_id) %>%
    select("Exp._ID", "Compound_ID", "AUC")

table2 <- head(AUC_2_rename)
kable(table2, align = "ccc")

```

### 3c. Experiment info
```{r Experiment info}

# Rename the columns in experiment_info file

experiment_info_rename <- experiment_info %>%
    rename("Exp._ID" = expt_id, "Date" = experiment_date, "Cells_per_well" = cells_per_well, "Cell_line_ID" = master_ccl_id) %>%
    mutate(Date = ymd(Date)) %>%
    select("Exp._ID", "Date", "Cell_line_ID", "Cells_per_well")

table3 <- head(experiment_info_rename)
kable(table3, align = "cccc")

```

### 3d. Cancer cell line info
```{r Cancer cell line info}

# Rename the columns in cancer_cell_line_info file

cancer_cell_line_rename <- cancer_cell_line %>%
    rename("Cell_line_name" = ccl_name, "Cancer_type" = cancer_type, "Cell_line_ID" = master_ccl_id)

table4 <- head(cancer_cell_line_rename)
kable(table4, align = "cll")

```

### 3e. Compound info
```{r Compound info}

# Rename the columns in compound_info file

compound_rename <- compound %>%
    rename("Compound_ID" = master_cpd_id, "Compound_name" = cpd_name, "Target_gene" = gene_symbol_of_protein_target)

table5 <- head(compound_rename)
kable(table5, align = "clc")

```
### 3f. Combine the data frames
```{r Combining data frames}

# Combine the AUC_1 and AUC_2 data frames into one - AUC_info

AUC_info <- rbind(AUC_1_rename, AUC_2_rename)

# Combine experiment_info with cancer_cell_line_info 

experiment_cancer_cell_line <- experiment_info_rename %>%
      full_join(cancer_cell_line_rename, by = "Cell_line_ID")

# Combine AUC_info and experiment_cancer_cell_line

AUC_exp_cancer <- AUC_info %>%
      full_join(experiment_cancer_cell_line, by = "Exp._ID")

# Add compound_info to experiment_cancer_cell_line

AUC_exp_cancer_compound <- AUC_exp_cancer  %>%
     full_join(compound_rename, by = "Compound_ID")

# Rename the column in the combined data frame

all_data <- AUC_exp_cancer_compound %>%
    rename("Cells/well" = "Cells_per_well")

# Rorder the columns in the combined data frame

all_data<- select(all_data, "Date", "Exp._ID", "Cell_line_name", "Cell_line_ID", "Cancer_type", "Cells/well", "Compound_name", "Compound_ID", "AUC", "Target_gene")

# Replace the NAs in Target gene column

target_gene <- all_data$`Target_gene` %>%
    replace_na("-")

final_data <- all_data %>%
    select(-'Target_gene') %>%
    mutate('Target_gene' = target_gene)

# I'm showing the final table in two parts for better visibility

table6 <- head(final_data[,1:6])
kable(table6, align = "lccclc")
    
table7 <- head(final_data[,7:10])
kable(table7, align = "lccc")

```

## 4. Questions

### 4a. Which cancer type has the lowest AUC values to the compound "vorinostat"?

```{r Vorinostat}

# Calculate mean AUC for vorinostat used alone or in combination with other drug

vorinostat_mean <- final_data %>%
    select("Cell_line_name", "Cell_line_ID", "Cancer_type", "Compound_name", "Compound_ID", "AUC") %>%
    separate(col = "Compound_name", into = c("Compound_1", "Compound_2")) %>%
    filter(Compound_1 == "vorinostat" | Compound_2 == "vorinostat") %>%
    group_by(Cancer_type) %>%
    summarize(AUC = round(mean(AUC), digits = 2)) %>%
    arrange(AUC)

table8 <- head(vorinostat_mean)
kable(table8, align = "lc", col.names = c("Cancer type", "Mean AUC"))

# Plot mean AUC for vorinostat used alone or in combination with other drug

vorinostat1 <- final_data %>%
    select("Cell_line_name", "Cell_line_ID", "Cancer_type", "Compound_name", "Compound_ID", "AUC") %>%
    separate(col = "Compound_name", into = c("Compound_1", "Compound_2")) %>%
    filter(Compound_1 == "vorinostat" | Compound_2 == "vorinostat")

ggplot(vorinostat1, aes(x = Cancer_type, y = AUC, fill = Cancer_type)) +

    geom_boxplot() +

    labs(title = "AUC values of different cancer types to vorinostat alone \nor in combination with other drug", x = "Cancer type", y = "AUC") +

    theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold", hjust = 0.5),
          plot.title.position = "plot",
          legend.position = "none",
          axis.title.x = element_text(colour="black", size = 12, face = "bold"),
          axis.title.y = element_text(colour="black", size = 12, face = "bold"),
          axis.ticks.x = element_line(size = 0.5),
          axis.text.y = element_text(colour="black", size = 10),
          panel.grid.major.x = element_blank(),
          panel.background = element_rect(fill = "white", color = "black")) +
    
    scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20), limits = c(0, 20)) +
    coord_flip()

# Calculate mean AUC for vorinostat used alone

vorinostat2_mean <- final_data %>%
    select("Cell_line_name", "Cell_line_ID", "Cancer_type", "Compound_name", "Compound_ID", "AUC") %>%
    separate(col = "Compound_name", into = c("Compound_1", "Compound_2")) %>%
    filter(Compound_ID == 56554) %>%
    group_by(Cancer_type) %>%
    summarize(AUC = round(mean(AUC), digits = 2)) %>%
    arrange(AUC)

table9 <- head(vorinostat2_mean)
kable(table9, align = "lc", col.names = c("Cancer type", "Mean AUC"))

# Plot mean AUC for vorinostat used alone

vorinostat2 <- final_data %>%
    select("Cell_line_name", "Cell_line_ID", "Cancer_type", "Compound_name", "Compound_ID", "AUC") %>%
    separate(col = "Compound_name", into = c("Compound_1", "Compound_2")) %>%
    filter(Compound_ID == 56554)

ggplot(vorinostat2, aes(x = Cancer_type, y = AUC, fill = Cancer_type)) +

    geom_boxplot() +

    labs(title = "AUC values of different cancer types to vorinostat alone", x = "Cancer type", y = "AUC") +

    theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold", hjust = 0.5),
          plot.title.position = "plot",
          legend.position = "none",
          axis.title.x = element_text(colour="black", size = 12, face = "bold"),
          axis.title.y = element_text(colour="black", size = 12, face = "bold"),
          axis.ticks.x = element_line(size = 0.5),
          axis.text.y = element_text(colour="black", size = 10),
          panel.grid.major.x = element_blank(),
          panel.background = element_rect(fill = "white", color = "black")) +
    
    scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20), limits = c(0, 20)) +
    coord_flip()

```
### _Vorinostat alone is most effective with cancers of autonomic ganglia. In combination with other drugs it is the most effective with cancers of haematopoietic and lymphoid tissues._

### 4b. Which compound is the prostate cancer cell line 22RV1 most sensitive to? 
###     (For 22RV1, which compound has the lowest AUC value?)

```{r 22RV1 prostate cancer}

# Calculate mean AUC for drugs tested with 22RV1 prostate cancer cell line

prostate_cancer <- final_data %>%
    select("Cell_line_name", "Cancer_type", "Compound_name", "Compound_ID", "AUC") %>%
    filter(Cell_line_name == "22RV1") %>%
    group_by(Compound_name) %>%
    summarize ("Mean_AUC" = round(mean(AUC), digits = 2)) %>%
    arrange(Mean_AUC)

table10 <- head(prostate_cancer)
kable(table10, align = "lc", col.names = c("Compound name", "Mean AUC"))

```
### _The prostate cell line 22RV1 is the most sensitive to leptomycin B._

### 4c. For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

```{r EGFR}

# Calculate mean AUC for 10 compounds targeting EGFR in breast cancer cell lines

EGFR_mean <- final_data %>%
    select("Cancer_type", "Compound_name", "Compound_ID", "AUC", "Target_gene") %>%
    mutate(EGFR = str_detect(Target_gene, "EGFR")) %>%
    filter(Cancer_type == "breast" & EGFR == TRUE) %>%
    group_by(Compound_name) %>%
    summarize("Mean_AUC" = round(mean(AUC), digits = 2)) %>%
    arrange(Mean_AUC)
 
table11 <- head(EGFR_mean)
kable(table11, align = "lc", col.names = c("Compound name", "Mean AUC"))

# Plot mean AUC for 10 compounds targeting EGFR in breast cancer cell lines

EGFR_plot <- final_data %>%
    select("Cancer_type", "Compound_name", "Compound_ID", "AUC", "Target_gene") %>%
    mutate(EGFR = str_detect(Target_gene, "EGFR")) %>%
    filter(Cancer_type == "breast" & EGFR == TRUE)

ggplot(EGFR_plot, aes(x = Compound_name, y = AUC, fill = Compound_name)) +

    geom_boxplot() +

    labs(title = "AUC values of 10 drugs targeting EGFR in breast cancer", x = "Drug", y = "AUC") +

    theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold", hjust = 0.5),
          plot.title.position = "plot",
          legend.position = "none",
          axis.title.x = element_text(colour="black", size = 12, face = "bold"),
          axis.title.y = element_text(colour="black", size = 12, face = "bold"),
          axis.ticks.x = element_line(size = 0.5),
          axis.text.y = element_text(colour="black", size = 10),
          panel.grid.major.x = element_blank(),
          panel.background = element_rect(fill = "white", color = "black")) +
    
    scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20), limits = c(0, 20)) +
    coord_flip()

```

### _Among the 10 drugs targeting EGFR gene in breast cancer cell lines, afatinib has the lowest average AUC._