---
title: "HW05"
author: "Ariana F. Tortolani"
date: "8/3/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PART A
Tidy the dad_mom data.

INSTRUCTIONS:
Clean this up so there are only 4 columns Tidy this data frame so that it adheres to the tidy data principles:
  Each variable must have its own column.
  Each observation must have its own row.
  Each value must have its own cell.

```{r load packages}
#load necessary packages
library(tidyverse)
library(knitr)
```

Here is an initial look at the data.
```{r dad_mom_initial}
#read in data
dad_mom_initial <- read.delim("dad_mom.txt")

#initial glimpse / visualization of data
glimpse(dad_mom_initial)
kable(dad_mom_initial, format = "markdown")
```

Tidy the data.
```{r dad_mom_tidy}
#tidy the data
dad_mom_tidy <- dad_mom_initial %>%
  #unite data together for dad and mom
  unite(Dad, name_dad, income_dad, sep = "_") %>% #unite all dad data
  unite(Mom, name_mom, income_mom, sep = "_") %>% #unite all mom data
  
  #gather data
  gather(Dad, Mom, -fam_id, key = "Parent", value = "Name_Income") %>% #creates Parent col 
  
  #separate name_income col
  separate(Name_Income, into = c("Name", "Income"), sep = "_") %>%
  
#make more readable
  #clean col names
  rename(Family = fam_id) %>%
  #sort by family id
  arrange(Family)

#view the tidy data
kable(dad_mom_tidy, format = "markdown")
```


## PART B

Join together the CTRP data and then answer the three questions presented below.

```{r load data}
#read in all the data
AUC_1 <- read.csv("CTRP_files/AUC_1.csv")
AUC_2 <- read.csv("CTRP_files/AUC_2.csv")
cancer_cell_line_info <- read.csv("CTRP_files/cancer_cell_line_info.csv")
compound_info <- read.csv("CTRP_files/compound_info.csv")
experiment_info <- read.csv("CTRP_files/experiment_info.csv")
```

to get an initial look at the data and see what variables each file contains
```{r view data}
glimpse(AUC_1)
glimpse(AUC_2)
glimpse(cancer_cell_line_info)
glimpse(compound_info)
glimpse(experiment_info)
```

join together the separated data frames into one complete data frame with all the information
```{r join data}
joined_CTRP_data <- AUC_1 %>%
  #combine AUC_1 and AUC_2 (stack on top of each other)
  bind_rows(AUC_2) %>%
  #combine experiment
  inner_join(experiment_info, by = c("experiment_id" = "expt_id")) %>%
  #combine cancer_cell_line)info
  inner_join(cancer_cell_line_info, by = "master_ccl_id") %>%
  #combine compound_info
  inner_join(compound_info, by = "master_cpd_id")

#view combined data frame
glimpse(joined_CTRP_data)
```

# Answering questions using the joined data frame:

1. Which cancer type has the lowest AUC values to the compound "vorinostat"?

ANSWER: From the graph below haematopoietic_and_lymphoid_tissue has the lowest AUC values to the compound vorinostat.
```{r Question 1}
#create new data frame to make plot
vorinostat_data <- joined_CTRP_data %>%
  select(area_under_curve, cancer_type, cpd_name) %>%
  filter(cpd_name == "vorinostat") %>%
  filter(cancer_type != "NA") %>%
  group_by(cancer_type)

#create plot
ggplot(vorinostat_data, aes(reorder(cancer_type, area_under_curve), area_under_curve)) +
  #use reorder to organize from lowest to highest value
  geom_boxplot() +
  
  #add labels
  labs(title = "Cancer Type Responsiveness to Vorinostat",
       x     = "Cancer Type",
       y     = "Area Under the Curve") +
  #make easier to read
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Which compound is the prostate cancer cell line 22RV1 most sensitive to? (For 22RV1, which compound has the lowest AUC value?)

ANSWER: When trying to answer this question i keep getting an error message that the ccl_name "22RV1" is not found so I am not sure if I created this data frame wrong or it just doesn't exist?
```{r Question 2}
#create new data frame to make plot
AUC_22RV1_data <- joined_CTRP_data %>%
  select(area_under_curve, cancer_type, cpd_name) %>%
#  filter(ccl_name == "22RV1") %>% #unexpected error, no data with name "22RV1"
  arrange(desc(area_under_curve))

```

3. For the 10 compounds that target EGFR, which of them has (on average) the lowest AUC values in the breast cancer cell lines?

ANSWER:  Based on the graph below the compound afatinib has the lowest AUC values.
```{r Question 3}
#create new data frame to make plot
EGFR_data <- joined_CTRP_data %>%
  select(area_under_curve, cancer_type, gene_symbol_of_protein_target, cpd_name) %>%
  filter(cancer_type == "breast") %>%
  filter(str_detect(gene_symbol_of_protein_target, "EGFR")) %>%
  group_by(cpd_name)

#create plot
ggplot(EGFR_data, aes(reorder(cpd_name, area_under_curve), area_under_curve)) +
  #use reorder to organize from lowest to highest value
  geom_boxplot() +

  #add labels
  labs(title = "EGFR targeting compounds AUC values in breast cancer",
       x     = "Compound Name",
       y     = "Area Under the Curve") +
  #make easier to read
  theme(axis.text.x = element_text(angle = 75, hjust = 1))
```

